---
title: "ST_512-homework6"
format: pdf
author: "Lindsay Dodson"
---

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(tidyverse)
```

## Question 1

```{r}
beer<- read.csv("beer_data.csv")
```

**a**

$H_0: \sigma^2_a=0$

$H_A: \sigma^2_a>0$

```{r}
MSa<- 170.91
MSe<- 0.72
dfa<- 5
dfe<-42
n<- 8

f_val<-MSa/MSe
1-pf(f_val, dfa, dfe)
```

The variation in sodium is due to brand as seen through p=2e-16\<0.05.

**b**

```{r}
beer|>
  group_by(brand) |>
  summarise("avg_sodium"=mean(sodium))
```

The mean sodium for Brand 1 is 23.6375 and Brand 2 is 10.6750.

**c**

$\hat{\sigma^2} = 0.72$

```{r}
(MSa-MSe)/n
```

$\hat{\sigma^2_\mu} = 21.27$

```{r}
21.27/(21.27+0.72)
```

97% of the variance in sodium in beer is associated with the brand of beer.

## Question 2

| source           | df  | SS   | MS     | F-va  |
|------------------|-----|------|--------|-------|
| rater            | 7   | 1641 | 234.43 | 5.019 |
| product          | 4   | 892  | 223    | 4.774 |
| interaction R\*P | 28  | 1308 | 46.71  | 6.36  |
| error            | 120 | 881  | 7.34   |       |

$\hat{\sigma^2}=7.34$

$\hat{\sigma^2_{\alpha\beta}}=\frac{46.71-7.34}{4}=9.843$

$\hat{\sigma^2_\alpha}=\frac{234.43-46.71}{5*4}=9.386$

$\hat{\sigma^2_\beta}=\frac{223-46.71}{8*4}=5.509$

## Question 3

**a**

| source      | df  | SS       | MS     | F-val  | p-val  |
|-------------|-----|----------|--------|--------|--------|
| location    | 1   | 1.30     | 1.30   | 0.0119 | 0.913  |
| player      | 9   | 3879.30  | 431.03 | 11.99  | 0.0005 |
| interaction | 9   | 323.67   | 35.96  | 0.328  | 0.965  |
| error       | 144 | 15787.29 | 109.63 |        |        |
| total       | 339 | 19991.56 |        |        |        |

```{r}
1-pf(0.0119,1,144)
1-pf(11.99,9,9)
1-pf(0.328,9,144)
```

**b**

There is an interaction between player and location as seen through $\sigma^2_{\alpha\beta}=35.96 > 0$

**c**

There is not a location main effect as seen through p=0.913>0.05.

**d**

There is a player main effect as seen through $\sigma^2_\beta= 431.03 > 0$

**e**

$\hat{\sigma^2_{\alpha\beta}}=\frac{35.96-109.63}{17}=-4.33$

$\hat{\sigma^2_\beta}=\frac{431.03-109.63}{17*2}=9.45$

$\hat{\sigma^2}=109.63$

## Question 4



